---
title: "Project 1"
author: "Ben Spanswick, Chase Henderson"
date: "October 17, 2018"
output: html_document
---

###Package Import

```{r, message=FALSE, warning=FALSE}
library(readr, quietly = TRUE)
library(tables, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(plyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(RColorBrewer, quietly = TRUE)
library(psych, quietly = TRUE)
library(highcharter, quietly = TRUE)
library(rhandsontable, quietly = TRUE)
library(xml2, quietly = TRUE)
library(naniar, quietly = TRUE)
library(pastecs, quietly = TRUE)
library(maps, quietly = TRUE)
library(maptools, quietly = TRUE)
library(sp, quietly = TRUE)
library(googleVis, quietly = TRUE)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Importing Data##

In the following code we are simply importing the beer and brewery data. We are importing data from the census in order to enrich our current datasets. 

```{r}
#Import Data

population <- read.table("https://people.sc.fsu.edu/~jburkardt/datasets/census/census_2010.txt")
names(population) <- c("State", "Population")
population$Population <- as.numeric(gsub(",", "", population$Population))
Beers <- read_csv("Beers.csv")
Breweries <- read_csv("Breweries.csv")
Breweries$State <- as.factor(Breweries$State)
Breweries <- as.data.frame(merge(Breweries, population, all = TRUE))
Breweries2 <- rename(Breweries, c("Brew_ID" = "Brewery_id", "Name" = "Brewery_Name"))


head(Breweries)

```


## How many breweries are present in each state?

We are able to provide a simple output, although not particuarly readable.

```{r}

##How many breweries in each state?
Breweries_Per_State <- with(Breweries, table(State))
BPS_DF <- as.data.frame(Breweries_Per_State)
BPS_DF <- as.data.frame(merge(BPS_DF, population))

BPS_DF$Per100k <- BPS_DF$Freq / (BPS_DF$Population/100000)
head(BPS_DF)

```

```

We are able to plot the data on a barchart to better visualize. 

```{r}
ggplot(BPS_DF, aes(reorder(State, Freq), Freq)) + geom_bar(stat = "identity") +
  geom_col(aes(fill = Freq)) +
  theme(plot.title = element_text(size=14, face="bold", hjust = .5), 
        axis.text.x = element_text(angle = 90, hjust = .5))+
  
  scale_fill_gradient2(low = "yellow", 
                       high = "red",
                       mid = "orange",
                       space = "Lab") + 
  
  labs(x = "States", y = "Number of Breweries") +
  
  ggtitle("Number of Breweries Per State")




```

We are also able to provide a heat map in order to represent the data geographically. 

```{r}

dat <- BPS_DF
names(dat)[1] <-'state.abb'
dat$states <- tolower(state.name[match(dat$state.abb,  state.abb)])

mapUSA <- map('state',  fill = TRUE,  plot = FALSE)
nms <- sapply(strsplit(mapUSA$names,  ':'),  function(x)x[1])
USApolygons <- map2SpatialPolygons(mapUSA,  IDs = nms,  CRS('+proj=longlat'))

idx <- match(unique(nms),  dat$states)
dat2 <- data.frame(value = dat$Freq[idx], state = unique(nms))
row.names(dat2) <- unique(nms)

USAsp <- SpatialPolygonsDataFrame(USApolygons,  data = dat2)

spplot(USAsp['value'], main = "Number of Breweries Per State (lower 48)")
```

Maybe number of breweries per state is a less than perfect metric. We will also examine number of breweries per 1000 residents of a state. 

```{r}

ggplot(BPS_DF, aes(reorder(State, Per100k), Per100k)) + geom_bar(stat = "identity") +
geom_col(aes(fill = Per100k)) +
  theme(plot.title = element_text(size=14, face="bold", hjust = .5), 
        axis.text.x = element_text(angle = 90, hjust = .5))+
  
  scale_fill_gradient2(low = "yellow", 
                       high = "red",
                       mid = "orange",
                       space = "Lab") + 
  
  labs(x = "States", y = "Number of Breweries Per 100k Residents") +
  
  ggtitle("Number of Breweries Per 100k Residents")

```


And provide a geographic representation of this data. 

```{r}
dat <- BPS_DF
names(dat)[1] <-'state.abb'
dat$states <- tolower(state.name[match(dat$state.abb,  state.abb)])

mapUSA <- map('state',  fill = TRUE,  plot = FALSE)
nms <- sapply(strsplit(mapUSA$names,  ':'),  function(x)x[1])
USApolygons <- map2SpatialPolygons(mapUSA,  IDs = nms,  CRS('+proj=longlat'))

idx <- match(unique(nms),  dat$states)
dat2 <- data.frame(value = dat$Per100k[idx], state = unique(nms))
row.names(dat2) <- unique(nms)

USAsp <- SpatialPolygonsDataFrame(USApolygons,  data = dat2)

spplot(USAsp['value'],main = "Number of Breweries Per 1000 Residents (lower 48)")

```




## Merge beer data with the breweries data and print the first 6 observations and the last six observations to check the merged file.

```{r}
Beer_Merge <- as.data.frame(merge(Beers, Breweries2, all = TRUE))
rhandsontable(head(Beer_Merge, n= 6))
rhandsontable(tail(Beer_Merge, n= 6))

```

##Reported number of NA's in each column

```{r}

rhandsontable(miss_var_summary(Beer_Merge))

```

We are also able to plot the missing variables per column. 

```{r}

gg_miss_var(Beer_Merge) 

```

##Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r}

ABV_IBU <- as.data.frame(aggregate(cbind(Beer_Merge$ABV, Beer_Merge$IBU) ~ Beer_Merge$State, Beer_Merge, median))

names(ABV_IBU) <- c("State", "ABV", "IBU")

rhandsontable(head(ABV_IBU))

```

Here we will use an interactive chart. Feel free to click and compare. 

```{r}
highchart() %>% 
  hc_yAxis_multiples(
    list(lineWidth = 3, lineColor='blue', title=list(text="ABV")),
    list(lineWidth = 3, lineColor="red", title=list(text="IBU"))
  ) %>% 
  hc_add_series(data = ABV_IBU$ABV, color='blue', type = "column") %>% 
  hc_add_series(data = ABV_IBU$IBU, color='red', type = "column", yAxis = 1) %>%
  hc_xAxis(categories = ABV_IBU$State, title = list(text = "State")) %>%
  hc_title(text = "<b>Median ABV and IBU by State</b>",
           margin = 20, align = "center",
           style = list(color="black", useHTML = TRUE))
```

##Which state has the maximum alcoholic (ABV) beer? 

```{r}

MaxABV <- Beer_Merge %>% 
    slice( which.max(ABV)) %>% 
    select(State, Name, ABV)

MaxIBU <- Beer_Merge %>% 
    slice( which.max(IBU)) %>% 
    select(State, Name, IBU)


rhandsontable(MaxABV, rowHeaders = NULL, trimWhitespace=TRUE)


```

##Which state has the most bitter (IBU) beer?

```{r}
rhandsontable(MaxIBU, rowHeaders = NULL, trimWhitespace=TRUE)
```


##Summary statistics for the ABV variable.

```{r}


sumStat <- as.data.frame(stat.desc(Beer_Merge$ABV))
names(sumStat) <- c("Summary")

rhandsontable(sumStat, rowHEaders=NULL, trimWhitespace=TRUE)

```

We can also provide a visual representation of ABV with a box plot. 

```{r}
ggplot(Beer_Merge, aes(x="", y=Beer_Merge$ABV)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4) + stat_summary(fun.y=mean, geom="point", shape=23, size=4) + labs(y = "ABV", x = "", title = "Summary of ABV")
```

##Summary Statistics for the IBU variable

```{r}
sumStat <- as.data.frame(stat.desc(Beer_Merge$IBU))
names(sumStat) <- c("Summary")

rhandsontable(sumStat, rowHEaders=NULL, trimWhitespace=TRUE)
```

And the visual representation...

```{r}

ggplot(Beer_Merge, aes(x="", y=Beer_Merge$IBU)) + 
  geom_boxplot(outlier.colour="Blue", outlier.shape=8,
                outlier.size=4) + stat_summary(fun.y=mean, geom="point", shape=23, size=4) + labs(y = "IBU", x = "", title = "Summary of IBU")

```

##Is there an apparent relationship between the bitterness of the beer and its alcoholic content?

It seems that from the scatterplot that there is a relationship. 

```{r}
Scatter_plot <- ggplot(Beer_Merge, aes(x = Beer_Merge$ABV, y= Beer_Merge$IBU)) + geom_point() + geom_smooth(method = "lm", se=TRUE, fullrange= TRUE, level= 0.95) + xlab("ABV") + ylab("IBU")

Scatter_plot
```

To confirm we can perform a one-way ANOVA. It appears that there is a significant relationship between the two. 

```{r}
res.aov <- aov(Beer_Merge$ABV ~ Beer_Merge$IBU, data = Beer_Merge)

summary(res.aov)
```


